Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-04 evolution.tex
cleaning up stuff
small improvement

  \section{Evolution Monitoring}

  When thinking about performance, 
  one can not avoid thinking about the performance evolution.

@@

  However, with the current configuration of the tool, it would be impossible for the maintainer to see the improvements resulting from the optimization.

  \ins{The latest STack Overflow Developer Survey: 88.4\% of **professional** developers use git.}  
  If assume that the code that they run is deployed using .git, then with an extra line of configuration they can allow \tool to find the git\footnote{\url{https://git-scm.com/}} folder of the deployed service and automatically detect the version of the project that is running:

@@

  \subsubsection*{Evolving Utilization}

  \Fref{fig:mv-util} shows the \perspective{Multi-Version Utilization} perspective. 
  It shows the utilization of all the tracked endpoints across versions.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\linewidth]{utilization-evolution}
      \caption{The Evolution of Endpoint Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

    \subsubsection*{Evolving Performance}

@@

  CHALLENGE: Supporting other types of version control... not real challenge, as we said 88percent of devs use git. 
  CHALLENGE 2: detecting minor versions which don't need to be tracked, since they didn't touch the performance of the system. E.g. a modification of the README, etc.

\todo{TODO: endpoint evolution }

\subsubsection*{Evolving Groups}
The limitation of the previous view is that it does not present the information also on a per version basis. To address this, a different visual perspective entitled \perspective{Multi-Version per-User Performance} can be defined. Figure \ref{fig:tuv} presents such a perspective by mapping the average execution time for a given user (lines) and given version (columns) on the area of the corresponding circle.

>>>>

  \section{Version-Aware Monitoring}

  The Stack Overflow Developer Survey from 2018 revealed the fact that 88.4\% of professional developers use git for version control. 
  Given that versioning using git is the main way in which the source code of services evolves, it is natural to monitor their progress also across versions.

@@

  % However, with the current configuration of the tool, it would be impossible for the maintainer to see the improvements resulting from the optimization.

  If assume that the code that they run is deployed using .git, then with an extra line of configuration they can allow \tool to find the git\footnote{\url{https://git-scm.com/}} folder of the deployed service and automatically detect the version of the project that is running:

@@

  \subsection*{Evolving Utilization}

  \Fref{fig:mv-util} shows the \perspective{Multi-Version API Utilization} perspective which presents the utilization of the tracked endpoints across versions. Because the different versions might be deployed for very different periods of time, at the intersection of an endpoint and a version the chart does not plot the absolute utilization of that endpoint in that version but rather the percentage of all the API calls that go to that endpoint. Otherwise a version that is deployed for many weeks would make all those deployed for several days invisible.

    \begin{figure}[h!]
      \centering
      \includegraphics[width=0.9\linewidth]{utilization-evolution}
      \caption{The Evolution of All the API Endpoints Utilization Across System Versions}
      \label{fig:mv-util}
    \end{figure}

  With this view, several patterns are visible:
  \begin{itemize}

    \item \epFeedItems is discontinued a little bit after \epTopArticles is being introduced

  \end{itemize}

  \subsection*{Evolving Performance}

@@

\subsection*{Evolving Groups}
  The limitation of the previous view is that it does not present the information also on a per version basis. To address this, a different visual perspective entitled \perspective{Multi-Version per-User Performance} can be defined. Figure \ref{fig:tuv} presents such a perspective by mapping the average execution time for a given user (lines) and given version (columns) on the area of the corresponding circle.

### 2018-04-04 performance.tex
done or the day
work in progress
cleaning up stuff
small improvement

%!TEX root=restructured.tex

\subsection{Service Performance}
\label{sec:perf}

@@

 \begin{figure}[!ht]
   \centering
   \includegraphics[width=0.95\linewidth]{endpoint_performance_}
   \caption{The response time (in ms) per monitored endpoint view allows for identifying performance variability and balancing issues}
   \label{fig:ep}
 \end{figure}

In the Zeeguu case study, one of the slowest endpoints, and one with the highest variability as shown in \Fref{fig:ep} is \epFeedItems: it retrieves a list of recommended articles for a given user. However, since a user can be subscribed to anything from one to three dozen article sources, and since the computation of the difficulty is personalized and it is slow, the variability in time among users is likely to be very large.

  From this view it became clear to the maintainer that four of the endpoints had very large variation in performance.   The most critical for the application and consequently the one optimized first was the \epTranslations endpoint which was part of an interactive loop in the reader applications that relied on the Zeeguu API. Moreover, cf. \Fref{fig:aeu} this endpoint is one of the most used in the system.

  \newpage
  \subsection{Automated Outlier Detection and Monitoring}

@@

  \begin{figure}[h!]
    \centering
    \includegraphics[width=\linewidth]{outlier-annotated}
    \caption{Automatically collected outlier information}
    \label{fig:figure1}
  \end{figure}

@@

>>>>

%!TEX root=paper.tex

\newpage
\subsection{Performance}
\label{sec:perf}

@@

 \begin{figure}[!ht]
   \centering
   \includegraphics[width=0.85\linewidth]{endpoint_performance_}
   \caption{The response time (in ms) per monitored endpoint view allows for identifying performance variability and balancing issues}
   \label{fig:ep}
 \end{figure}

In the Zeeguu case study, one of the slowest endpoints, and one with the highest variability as shown in \Fref{fig:ep} is \epFeedItemsColor which retrieves a list of recommended articles for a given user. However, since a user can be subscribed to anything from one to three dozen article sources, and since the computation of the difficulty is personalized and it is slow, the variability in time among users is likely to be very large.

From this view it became clear to the maintainer that four of the endpoints had very large variation in performance.   The most critical for the application and consequently the one optimized first was \epTranslationsColor endpoint which was part of an interactive loop in the reader applications that relied on the Zeeguu API. Moreover, cf. \Fref{fig:aeu} this endpoint is one of the most used in the system.

  \subsection{Automated Outlier Detection and Monitoring}

@@

  \begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\linewidth]{outlier-annotated}
    \caption{Automatically collected outlier information}
    \label{fig:figure1}
  \end{figure}

@@

\lesson{We need more advanced outlier analysis tools... Instead of manually searching in the page to build stats about Google vs. Microsoft}

### 2018-04-04 utilization.tex
work in progress
.
smarter table
small improvement

\newpage
\subsection{Service Utilization}
\label{sec:util}

@@

  \begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{basicest-utilization}
  \caption{....}
  \label{fig:basicest}
  \end{figure}

@@

  \item {\bf Frequent Patterns of System Usage}. The two most used endpoints stand for the two main activities in the system:

@@

    \item \epTranslations is an indicator of the amount of foreign language reading the users are doing

    \epOutcome is an indicator of the amount of foreign vocabulary practice the users are doing.

@@

  \item {\bf Sudden Increase of Endpoint Usage}. One endpoint (\epUserActivity) has been disproportionately been used in the last day and last week; much more than before.

  \item {\bf Possibly Discontinued Endpoint Usage}. One endpoint (\epFeedItems) has not been used in the last 7 days

@@

The table view is limited, and the one day and one week periods are conveniently but arbitrarily selected. For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily Utilization} perspective on endpoint utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, a day when the API had more than 12.000 hits.

@@

Another utilization perspective is the \perspective{Hourly Utilization} in which  the \tool can highlight {\em cyclic patterns of usage per hour of day} by means of a heatmap, as in \Fref{fig:dp}.

>>>>

\newpage
\subsection{Utilization}
\label{sec:util}

@@

  \begin{figure}[h!]
  \centering
  \includegraphics[width=\linewidth]{utilization-table}
  \caption{....}
  \label{fig:basicest}
  \end{figure}

@@

  \item {\bf Application Specific Patterns of System Usage}. The two most used endpoints in the figure stand for the two main activities in the system:

@@

    \item \epTranslationsColor is an indicator of the amount of foreign language reading the users are doing

    \item \epOutcomeColor is an indicator of the amount of foreign vocabulary practice the users are doing.

@@

  \item {\bf Sudden Increase of Endpoint Usage}. One endpoint (\epUserActivityColor) has disproportionately been used in the last day and last week.

  \item {\bf Possibly Discontinued Endpoint Usage}. One endpoint (\epFeedItemsColor) has not been used in the last 7 days. This might be an endpoint that is deprecated.

@@

The table view is limited, and the one day and one week periods are conveniently but arbitrarily selected. For a more detailed evolution of utilization, Figure \ref{fig:aeu} shows the \perspective{Daily API Utilization} perspective on endpoint utilization that \tool provides: a stacked bar chart of the number of hits to various endpoints grouped by day\footnote{Endpoint colors are the same in different views}. Figure~\ref{fig:aeu} in particular shows a peak utilization, a day when the API had more than 12.000 hits.

@@

Another utilization perspective is the \perspective{Hourly API Utilization} in which  the \tool can highlight {\em cyclic patterns of usage per hour of day} by means of a heatmap, as in \Fref{fig:dp}.
